---
title: "PM566 Final Project"
author: "CJ"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

<br>

# Introduction

Force production asymmetry between limbs is often used as a predictor of future injury, with asymmetries greater than 10-15% being considered at risk (1). Increased symmetry as a result of rehabilitation can also be used as a marker of success (1). Although used as a predictor of future injury, especially in ACL rehab patients (2), the effect of interlimb force production asymmetry is less clear as it relates to performance. A systematic review (1) evaluating the effects of inter-limb asymmetries on physical and sports performance found a high rate of asymmetry across a range of physical quantities. While not a consistent finding, they also concluded that asymmetry has a detrimental effect on performance. Single leg and horizontal jumps have been shown to detect asymmetries, but their effect on Change of Direction performance appears inconclusive. Some studies have even shown a positive impact of asymmetry in cycling (3) and no effect in swimming (4). During bilateral jumping tasks, it has been shown that a significant difference in the asymmetry between landing and take off phases exists (5).

The counter-movement jump (CMJ) is divided into 5 phases with different biomechanical demands: Unweighing, Braking, Propulsion, Flight, and Landing. This project aimed to calculate the average asymmetry between limbs during each phase, as well as visualize the change in asymmetry across the course of the jump. By relating the asymmetry score to the phase being performed, corrective exercise mimicking the demands of that phase may be implemented in an attempt to restore symmetry.

# Methods

Data from a single CMJ performed by a 28 year old recreational hockey player was collected on a Vault force plate at a sampling rate of 1000 Hz at the USC Clinical Exercise Research Center on September 2, 2022. The data was exported to a csv file, and initially included 3136 observations and 22 variables. After importing the data in R using the function read.csv, the combined function mean(is.na) was used to confirm no missing data for the variables representing the force produced in the Z direction for the left and right legs.


<br>

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}

library(data.table)
library(tidyverse)
library(dplyr)
library(knitr)

# Initialize code chunk options
rm(list=ls())

library(ggplot2)
library(plotly)
library(readr)

# Initialize code chunk options

opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```





```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}

.code-r-small { /* Code block */
  font-size: 10px;
}
```

<br>

## First step

First I will source any necessary code, e.g. `process_covid_data.R`:

```{r load-data}
data <-read.csv("CJ CMJ-Countermovement Jump-2022.09.02-10.33.20-Trial1 - CJ CMJ-Countermovement Jump-2022.09.02-10.33.20-Trial1.csv",
                skip=9)
mean(is.na(data$Z.Left))
mean(is.na(data$Z.Right))
```

<br>

## Second step

Then I will add some code to create the **plotly** figures

Note: You need to name each code chunk, like this one: `plot1`

```{r plot1, class.source="code-r-small"}

# p1_scatter <- cv_states_today %>% 
#   plot_ly(x = ~pop_density, y = ~deathsper100k,
#           type = 'scatter', mode = 'markers', color = ~state,
#           size = ~population, sizes = c(5, 70), marker = list(sizemode='diameter', opacity=0.5),
#           hoverinfo = 'text',
#           text = ~paste( paste(state, ":", sep=""), paste(" Cases per 100k: ", per100k, sep="") , paste(" Deaths per 100k: ",
#                         deathsper100k, sep=""), sep = "<br>")) %>%
#   layout(title = "Population-normalized COVID-19 deaths vs. population density",
#                   yaxis = list(title = "Deaths per 100k"), xaxis = list(title = "Population Density"),
#          hovermode = "compare")
# 
# # filter out "District of Columbia"
# cv_states_today_scatter <- cv_states_today %>% filter(state!="District of Columbia")
# 
# p2_scatter <- cv_states_today_scatter %>% 
#   plot_ly(x = ~pop_density, y = ~deathsper100k,
#           type = 'scatter', mode = 'markers', color = ~state,
#           size = ~population, sizes = c(5, 70), marker = list(sizemode='diameter', opacity=0.5),
#           hoverinfo = 'text',
#           text = ~paste( paste(state, ":", sep=""), paste(" Cases per 100k: ", per100k, sep="") , paste(" Deaths per 100k: ",
#                         deathsper100k, sep=""), sep = "<br>")) %>%
#   layout(title = "Population-normalized COVID-19 deaths vs. population density",
#                   yaxis = list(title = "Deaths per 100k"), xaxis = list(title = "Population Density"),
#          hovermode = "compare")
```

<br>

## Third step {.tabset}

Create tabs to display each figure

### Scatterplot: with DC

```{r p1}
# p1_scatter
```

### Scatterplot: without DC

```{r p2}
# p2_scatter
```

## {-}

<br>

Done!

<br>
<br>