---
title: "Limb Asymmetry During a Single Counter-Movement Jump"
author: "CJ"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

<br>



[Click here](https://github.com/Cgeeza/cgeeza.github.io/blob/main/midterm--2-.pdf) to download my report.

Force production asymmetry between limbs is often used as a predictor of future injury, with asymmetries greater than 10-15% being considered at risk. Increased symmetry as a result of rehabilitation can also be used as a marker of success (1).

The counter-movement jump (CMJ) is divided into 5 phases with different biomechanical demands: Unweighing, Braking, Propulsion, Flight, and Landing. This project aimed to calculate the average asymmetry between limbs during each phase, as well as visualize the change in asymmetry across the course of the jump. By relating the asymmetry score to the phase being performed, corrective exercise mimicking the demands of that phase may be implemented in an attempt to restore symmetry.

Data from a single CMJ performed by a 28 year old recreational hockey player was collected on a Vault force plate at a sampling rate of 1000 Hz at the USC Clinical Exercise Research Center on September 2, 2022.




```{r, echo=FALSE}
rm(list=ls())

library(ggplot2)
library(plotly)
library(readr)

data <-read.csv("CJ CMJ-Countermovement Jump-2022.09.02-10.33.20-Trial1 - CJ CMJ-Countermovement Jump-2022.09.02-10.33.20-Trial1.csv",
                skip=9)

```



```{r, echo=FALSE}
# Change time to normalize to 0
data$Time <- data$Time - min(data$Time)
```



```{r, echo=FALSE}
T <- nrow(data)
# re-arrange data to be able to plot left leg against right leg
data_tidy <- data.frame(Time=data$Time, Force=data$Z.Left, Leg="Left")

data_tidy <- rbind(data_tidy, data.frame(Time=data$Time, Force=data$Z.Right,
                                         Leg="Right"))

# Create column for difference between legs
data$difference <- data$Z.Right - data$Z.Left



plot_both <- ggplot(data_tidy, aes(x=Time, y=Force, color=Leg)) + geom_line() +
  ylab("Force (Newtons)") + xlab("Time (seconds)") + ggtitle("Force Produced in Right and Left Leg over Time")
```





```{r, echo=FALSE}
# Find standing phase: acceleration 0
# Looks like rows 1 - 592
# data[2100:2600, c("Time", "Acceleration")]
standing_inds <- 1:592

# Find unweighing phase: acceleration negative
# Rows 593 - 973
unweighting_inds <- 593:973



# Find breaking phase: acceleration positive
# Rows 974 - 1374
braking_inds <- 974:1374

plot_diff <- ggplot(data[braking_inds, ], aes(x=Time, y=difference)) + geom_point() +
  ylab("Differences in Forces (Right Leg - Left) (Newtons)") + xlab("Time (seconds)") + 
  ggtitle("Difference in force production between limbs during Braking Phase")

# Find propulsion phase: acceleration peaks and starts declining
# Rows 1375 - 1601
propulsion_inds <- 1375:1601

# In the air: acceleration -9.8

# Landing rows: 2155 - T
landing_inds <- 2155:T
```



```{r, echo=FALSE}
# Calculate the percentage difference between force from left and right legs
# in the z direction at each time
data$pct_diff <- (data$Z.Left - data$Z.Right)/pmax(data$Z.Left, data$Z.Right)*100
data$abs_pct_diff <- abs(data$Z.Left - data$Z.Right)/pmax(data$Z.Left, data$Z.Right)*100

# Average absolute asymmetry during standing phase:
avg_standing_asymmetry <- mean(data[standing_inds, "abs_pct_diff"])

# Average absolute asymmetry during unweighting phase:
avg_unweighting_asymmetry <- mean(data[unweighting_inds, "abs_pct_diff"])

# Average absolute asymmetry during breaking phase:
avg_braking_asymmetry <- mean(data[braking_inds, "abs_pct_diff"])

# Average absolute asymmetry during propulsion phase:
avg_propulsion_asymmetry <- mean(data[propulsion_inds, "abs_pct_diff"])

# Average absolute asymmetry during braking phase
avg_landing_asymmetry <- mean(data[landing_inds, "abs_pct_diff"])
```


```{r, echo=FALSE}
df_plot <- data[c(standing_inds, unweighting_inds, braking_inds,
                  propulsion_inds), ]

plot_pct_diff <- ggplot(df_plot, aes(x=Time, y= pct_diff)) + geom_point() +
  geom_hline(yintercept=10, color="red", linetype = "dashed") +
  geom_hline(yintercept=-10, color="red", linetype = "dashed") +
  ylab("Percent Difference") + ggtitle("Percent Asymmetry in Leg Force Over Time (Before Landing)")

landing_plot <- data[landing_inds,]

landing_plot_pct_diff <- ggplot(landing_plot, aes(x=Time, y= pct_diff)) + geom_point() +
  geom_hline(yintercept=10, color="red", linetype = "dashed") +
  geom_hline(yintercept=-10, color="red", linetype = "dashed") +
  ylab("Percent Difference") + ggtitle("Percent Asymmetry in Leg Force Over Time During Landing")
```



```{r, echo=FALSE}
ggplotly(plot_diff)


ggplotly(plot_both)

ggplotly(plot_pct_diff)

ggplotly(landing_plot_pct_diff)
```





The subject displayed average asymmetries greater than 10% during the Standing(11.63 %), Unweighing (11.45%), and Landing phases (17.85%). Interestingly, the subject's asymmetry decreases as they transition from passive phases (Standing and Unweighing) to power absorption and generation phases (Braking and Propulsion). The large difference between takeoff and landing symmetry coincide with results from Cone et al. Based on these results, the subject should use exercise strategies that target static balance and landing mechanics in order to reduce risk for future injury.

